---
title: "My Final Project -Dynamic"
runtime: shiny
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
    social: menu
    theme: united
# run rsconnect::deployApp()
---

```{r global, include=FALSE}
library(flexdashboard)
library(rtweet)
library(leaflet)
library(quanteda)
library(quanteda.textstats)
library(quanteda.textplots)
library(igraph)
library(visNetwork)
library(scales)
library(highcharter)
library(lubridate)
library(reshape2)
library(dplyr)
library(syuzhet) 
library(stringr)
library(tidyr)
library(DT)

load(file="#WorldCup2022_tweets.rda")
load(file="sentiments_#WorldCup2022_tweets.rda")

```

Inputs {.sidebar}
====================================================================

```{r}
h5("Data updated:", Sys.Date())

checkboxGroupInput("sentiment1", 
                   label = (helpText(h5("Choose sentiments"))),
                    choices = list("Anger" = "anger", 
                    "Anticipation" = "anticipation",
                    "Disgust" = "disgust",
                    "Fear" = "fear",
                    "Joy" = "joy", 
                    "Sadness" = "sadness", 
                    "Surprise" = "surprise",
                    "Trust" = "trust"),
                   selected = list("surprise","trust","anger"))

selectInput("language", label = (helpText(h5(strong("Choose a language")))),
            choices = list("English" = "en",
                           "French" = "fr",
                           "Spanish" = "es",
                           "German" = "de",
                           "Chinese" = "zh",
                           "Arabic" = "ar",
                           "Russian" = "ru",
                           "Turkish" = "tr"),
            selected = "en"
)


```

Sentiments {data-orientation=scroll data-icon="fa-bar-chart"}
=======================================================================

### 
###

### **Introducing Sentiment Analysis**

The sentiment detection algorithm detects *anger*, *sadness*, *joy*, *trust*, *surprise*, *fear*, *anticipation*, and *disgust*. 

### Row 1 {data-height=110}
-----------------------------------------------------------------------

```{r}

senti_aggregated$date_label <- as.Date(senti_aggregated$date_label)

output$chart <- renderHighchart({

  highchart() %>%
    hc_add_series(data= senti_aggregated[senti_aggregated$variable %in% input$sentiment1,],"line", hcaes(x = date_label, y = value,                                                                 group=variable)) %>%
hc_xAxis(type = "datetime")

})

highchartOutput('chart')

```

Hashtags {data-orientation=scroll data-icon="fa-bar-chart"}
=======================================================================

### **Introducing hashtag networks by language**

Hashtag is defined as xxxxx.

```{r}
renderPlot({
  corpus <- corpus(tweets[tweets$lang == input$language,],docid_field = "id_str",text_field = "full_text")
  
  tweets_dfm <- tokens(corpus, remove_punct = TRUE, remove_numbers = TRUE, remove_symbols = TRUE, remove_url=TRUE) %>%  
  dfm()
  
  tag_dfm <- dfm_select(tweets_dfm, pattern = "#*")
  toptag <- names(topfeatures(tag_dfm, 100))
  tag_fcm <- fcm(tag_dfm)
  topgat_fcm <- fcm_select(tag_fcm, pattern = toptag)
  textplot_network(topgat_fcm, min_freq = 0.1, edge_alpha = 0.5, edge_size = 1, edge_color = "orange")

})

```

Map {data-orientation=scroll data-icon="fa-bar-chart"}
=======================================================================

### **Introducing geomap by language**

This is a map. Explain the map here...

```{r}

renderLeaflet({
  geocodes <- lat_lng(tweets[tweets$lang == input$language,])
  geocodes <- geocodes[!is.na(geocodes$lat),]

  map <- leaflet(data = geocodes) %>% 
  addTiles() %>%
  setView(lng = 39, lat = 34, zoom = 2) %>% 
  addMarkers(lng = ~lng, lat = ~lat,popup = ~ as.character(text)) %>% 
  addProviderTiles(providers$Esri.NatGeoWorldMap) %>% #select basemap template
  addCircleMarkers(stroke = FALSE, fillOpacity = 0.2) 
  
  map

})


```

